Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-04 Manuscript/manuscript.tex
Major change in order parameter changes plot -- split NaCl and CaCl2 plot to two separate figures as it follows the story better (and looks nicer, too)

The headgroup order parameter changes with increasing amount of 
cationic surfactant dihexadecyldimethylammonium bromide 
is compared between experiments~\cite{scherer89} and simulations in
Fig. \ref{OrderParameterCHANGESsurf}. 
The observed order parameter decrease is simulations and experiments
can be approximated to be linear at least with the mole fractions
below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
The slope is, however, too steep in Lipid14 model indicating that 
the head group order parameters are too sensitive to a bound charge.
The ECC-lipids model gives a slope in very good agreement with experiments
for the $\alpha$ segment, while the slope is slightly
underestimated for the $\beta$ segment.
%This has to be taken into account when interpreting the
%head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
%observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
%The present comparison of head group order parameter changes 
%suggests that the observed overestimated response of all models to 
%increasing CaCl$_2$ concentration in \cite{catte16} 
%can be explained at least in part by a high sensitivity of the head group order parameter response
%to the bound charge.
\todo{SAMULI: We could calculate the slopes from simulations, but I am not sure if
we would actually learn anything useful from this.}

@@

\begin{figure*}[tbp]
  \centering
  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2.eps}
  \caption{\label{OrderParameterCHANGESnewMODELS}
    Changes of head group order parameters of POPC bilayer as a function of NaCl and CaCl$_2$ concentrations
    are shown from simulations with different force fields together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
\end{figure*}

The binding affinity of aqueous cations in lipid 
bilayers can be measured by using the headgroup 
order parameters, because they decrease proportionally  
to the bound positive charge~\cite{seelig87,catte16}. 
The headgroup order parameter responses 
to aqueous NaCl and CaCl$_2$ concentrations 
are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}
from experiments (DPPC \cite{akutsu81} 
and POPC \cite{altenbach84}) and different simulation models for POPC.

@@

Negligible changes of the headgroup order parameters are measured
with submolar concentrations of NaCl due to the very low affinity of Na$^+$
in PC bilayers \cite{akutsu81}. While Na$^+$ binding and the related 
headgroup order parameter changes were overestimated in
almost all the available simulation models, the low affinity
and negligible order parameter changes were reproduced by Lipid14 
model when simulated with \AA{}qvist ions~\cite{catte16}.
However, the same combination of force field parameters
overestimated the headgroup order parameter response to CaCl$_2$ 
concentration, which was the case also in all other models tested in
Ref.~\citenum{catte16}.
Using ion model by Dang et al.~\cite{smith94,chang1999,dang2006} or
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
with more realistic bulk behaviour did not improve the results
for the CaCl$_2$ interactions with Lipid14 model, as seen 
in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
The results support the conclusion of the previous work \cite{catte16}
that improvements also in lipid models are needed to 
correctly describe divalent cation binding in PC bilayers.

Significant improvement can be achieved by using the ECC
approach also for lipids. The headgroup order parameter changes
as a function of CaCl$_2$ concentration from ECC-lipid model with ECC-ions
show a good agreement with experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
As discussed in previous section, the model gives also
a good agreement with experiments for
the headgroup response to bound charge.
Thus, the model can used for more detailed analysis of the
binding affinity.

@@

>>>>

The headgroup order parameter changes with increasing amount of 
cationic surfactant dihexadecyldimethylammonium bromide 
is compared between experiments~\cite{scherer89} and simulations in
Fig. \ref{OrderParameterCHANGESsurf}. 
The observed order parameter decrease is simulations and experiments
can be approximated to be linear at least with the mole fractions
below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
The slope is, however, too steep in Lipid14 model indicating that 
the head group order parameters are too sensitive to a bound charge.
The ECC-lipids model gives a slope in very good agreement with experiments
for the $\alpha$ segment, while the slope is slightly
underestimated for the $\beta$ segment.
%This has to be taken into account when interpreting the
%head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
%observed in  Ref. \citenum{catte16}, see also Fig. \ref{fig:delta_ordPar_NaCl}.
%The present comparison of head group order parameter changes 
%suggests that the observed overestimated response of all models to 
%increasing CaCl$_2$ concentration in \cite{catte16} 
%can be explained at least in part by a high sensitivity of the head group order parameter response
%to the bound charge.
\todo{SAMULI: We could calculate the slopes from simulations, but I am not sure if
we would actually learn anything useful from this.}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
  \caption{\label{fig:delta_ordPar_NaCl}
    Changes of head group order parameters of POPC bilayer as a function of NaCl concentrations
    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
\end{figure}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
\end{figure}

The binding affinity of aqueous cations in lipid 
bilayers can be measured by using the headgroup 
order parameters, because they decrease proportionally  
to the bound positive charge~\cite{seelig87,catte16}. 
The headgroup order parameter responses 
to aqueous NaCl and CaCl$_2$ concentrations 
are shown in Fig. \ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}
from experiments (DPPC \cite{akutsu81} 
and POPC \cite{altenbach84}) and different simulation models for POPC.

@@

Negligible changes of the headgroup order parameters are measured
with submolar concentrations of NaCl due to the very low affinity of Na$^+$
in PC bilayers \cite{akutsu81}. While Na$^+$ binding and the related 
headgroup order parameter changes were overestimated in
almost all the available simulation models, the low affinity
and negligible order parameter changes were reproduced by Lipid14 
model when simulated with \AA{}qvist ions~\cite{catte16}.
However, the same combination of force field parameters
overestimated the headgroup order parameter response to CaCl$_2$ 
concentration, which was the case also in all other models tested in
Ref.~\citenum{catte16}.
Using ion model by Dang et al.~\cite{smith94,chang1999,dang2006} or
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
with more realistic bulk behaviour did not improve the results
for the CaCl$_2$ interactions with Lipid14 model, as seen 
in Figs. \ref{fig:delta_ordPar_CaCl} and S\ref{fig:ordPars_waterModels} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
The results support the conclusion of the previous work \cite{catte16}
that improvements also in lipid models are needed to 
correctly describe divalent cation binding in PC bilayers.

Significant improvement can be achieved by using the ECC
approach also for lipids. The headgroup order parameter changes
as a function of CaCl$_2$ concentration from ECC-lipid model with ECC-ions
show a good agreement with experiments in Fig. \ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}.
As discussed in previous section, the model gives also
a good agreement with experiments for
the headgroup response to bound charge.
Thus, the model can used for more detailed analysis of the
binding affinity.

@@

\begin{figure*}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
  \caption{\label{fig:ordPars_waterModels}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields and water models together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
  }
\end{figure*}
